---
output: Cluster Analysis
  author: Kafayat Alobaloke
---

```{r}
library(factoextra)
library(corrplot)
```

```{r}
data <- read.csv('leaf.csv')
```

```{r}
# Summary of the data
summary(data)
```

# Str data gives the type in which data is stored, this shows its treating Date as Factor so that needs to ve convereted to Date format
```{r}
# Understanding the datatype of dataset
str(data)
```

In K-means Algorithm we have to define the number of Cluster which represent with k
fortunately in R, We have some interesting functions for selecting k.
```{r}
fviz_nbclust(data, kmeans, method = "wss")
# method can be = 'gap_stat', 'wss', 'silhouette'
fviz_nbclust(data, kmeans, method = "gap_stat")
fviz_nbclust(data, kmeans, method = "silhouette")
```

```{r}
km1 <- kmeans(data, 2)
km1
```

All method('wss', 'silhouette', and gap_stat) show us it's better choose k = 2
The function fviz_cluster() [factoextra package] can be used to easily visualize k-means clusters. It takes k-means results and the original data as arguments. In the resulting plot, observations are represented by points, using principal components if the number of variables is greater than 2
```{r}
fviz_cluster(km1, data = data,
  palette = c("#2E9FDF", "#00AFBB"),
  ellipse.type = "euclid", # Concentration ellipse
  star.plot = TRUE, # Add segments from centroids to items
  repel = TRUE, # Avoid label overplotting (slow)
  ggtheme = theme_minimal())
```

```{r}
km2 <- kmeans(data, 3)
km2
```

```{r}
fviz_cluster(km2, data = data,
  palette = c("#2E9FDF", "#00AFBB", "#ADFFBB"),
  ellipse.type = "euclid", # Concentration ellipse
  star.plot = TRUE, # Add segments from centroids to items
  repel = TRUE, # Avoid label overplotting (slow)
  ggtheme = theme_minimal())
```

```{r}
# Splitting data into training and testing with 60 and 40 percent respectively
index <- sample(2, nrow(data), replace=TRUE, prob = c(0.60, 0.40))
traindata <- data[index==1, ]
testdata <- data[index==2, ]
```

```{r}
plot_ <- function(i){
        cluster <- kmeans(traindata, centers = i, nstart = 25)
        plot_list <- fviz_cluster(cluster, geom = "point", data = traindata) + ggtitle(paste("k = ", i)) + theme_minimal()
        return(plot_list)
  }
    

library(gridExtra)
grid.arrange(plot_(2),plot_(3),plot_(4),plot_(5), nrow = 2) + theme_minimal()
```

```{r}
plot_ <- function(i){
        cluster <- kmeans(testdata, centers = i, nstart = 25)
        plot_list <- fviz_cluster(cluster, geom = "point", data = testdata) + ggtitle(paste("k = ", i)) + theme_minimal()
        return(plot_list)
  }
    

library(gridExtra)
grid.arrange(plot_(2),plot_(3),plot_(4),plot_(5), nrow = 2) + theme_minimal()
```

```{r}
leaf <- data[,-16]
leaf
```

```{r}
library(RColorBrewer)
palette(alpha(brewer.pal(9,'Set1'), 0.5))
plot(testdata, col=leaf$Specimen, pch=16, bty="n") + theme_minimal()
```

```{r}
# Converting the integer data to factor
leaf$Species <- as.factor(leaf$Species)
leaf$SpecimenNumber <- as.factor(leaf$SpecimenNumber)
```

```{r}
summary(leaf)
```

```{r}
ggplot(leaf, aes(x=SpecimenNumber, fill = SpecimenNumber)) + 
  theme_bw()+
  geom_bar()+
  labs(x = "SpecimenNumber", y = "Data Count")
```

```{r}
corr_mat <- cor(leaf[,3:ncol(leaf)])
corrplot(corr_mat, order = "hclust", tl.cex = 1, addrect = 8)
```

```{r}
ggplot(leaf, aes(x = SpecimenNumber, fill = SpecimenNumber)) +
  geom_density(alpha=0.5) +
  scale_fill_discrete(name = "Leaf Classification")
```
